---
title: "Airbnb Case Study"
author: "YC"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r getready, echo=FALSE, message=FALSE}
setwd("C:/Users/YC/portfolio/case_study/airbnb")
#install.packages("sqldf")
library("sqldf")
library("ggplot2")
contacts <- read.csv(file="contacts.csv", header=TRUE, sep=",")
users <- read.csv(file="users.csv", header=TRUE, sep=",")
listings <- read.csv(file="listings.csv", header=TRUE, sep=",")
```

```{r getreadey2, echo=FALSE}
#replacing negative number of reviews with 0
listings$total_reviews[listings$total_reviews < 0] <- 0
```

##### Airbnb is a two sided marketplace which matches guests to hosts. The booking flow at Airbnb is as follows: a guest finds an available room (listing) that they like, and then they contact the host. Once the guest finds a listing they are interested in, there are three ways to send the host an inquiry: ‘contact_me’, ‘book_it’, or ‘instant_book’ (detailed at the bottom of this document). Upon receiving the inquiry, the host can then decide whether or not to accept the request (for ‘contact_me’ and ‘book_it’ methods -- `instant_book` is auto-accepted). One of our goals at Airbnb is to increase bookings on our platform. #####

```{r listings1}
l1 <- sqldf("select room_type, count(distinct id_listing_anon) as count_uniq_listing_id
            from listings l
            group by room_type
            order by count_uniq_listing_id desc
            ")
#Room Type
l1
l1$count_uniq_listing_id / sum(l1$count_uniq_listing_id) * 100

l2 <- sqldf("select listing_neighborhood, count(distinct id_listing_anon) as count_uniq_listing_id
            from listings l
            group by listing_neighborhood
            order by count_uniq_listing_id desc
            ")
l2
#Neighborhood
(l2$count_uniq_listing_id / sum(l2$count_uniq_listing_id) * 100)[1:10]


#
l3 <- sqldf("select room_type, avg(total_reviews) as avg_review_count
            from listings l
            group by room_type
            order by avg_review_count desc
            ")



#
qplot(total_reviews, data = listings, geom = "histogram", fill = room_type)

```
```{r listings2, fig.height=10}
qplot(room_type, total_reviews, data = listings, 
      geom= "boxplot", fill = room_type)
```


```{r users, fig.height=13}
u1 <- sqldf("select country, count(distinct id_user_anon) as count_uniq_users
            from users u
            group by country
            order by count_uniq_users desc
            ")
u1[1:10,]
u2 <- u1$count_uniq_users / sum(u1$count_uniq_users) * 100
u2[1:10]
hist(users$words_in_user_profile)
```
