---
title: "Simple Inference"
author: "YC"
output: html_document
---


```{r getready, echo=FALSE}
#install.packages("HSAUR")
#library("HSAUR")
data("roomwidth", package = "HSAUR")
#Note that the `echo = FALSE` Prevents code from printing
```

## Estimating the width of a room
*Data from HSAUR*

```{r rw_preview, echo=FALSE}
roomwidth[43:47,]
```

The data shown in the above table, as you can notice has estimates of room's width in both meters and feet. Each estimate was recorded from multiple individuals who were in a country where 'foot' was used predominantly over 'meter'.

If we convert meters to feet by multiplying each of them by a scalar factor of 3.28 then we would like to test the hypothesis that the mean of the population of ‘metre’ estimates is equal to the mean of the population of ‘feet’ estimates. We shall do this first by using an independent samples t-test, but first let's check the normality and equal variance assumptions.

```{r meter_conversion}
convert <- ifelse(roomwidth$unit == "feet", 1, 3.28)
```

Now some summary stats
```{r rw_summary}
tapply(roomwidth$width * convert, roomwidth$unit, summary)
```

standard deviation
```{r rw_summary_sd}
tapply(roomwidth$width * convert, roomwidth$unit, sd)
```

Now examine the data with some plots
```{r rw_plots}
layout(matrix(c(1, 2, 1, 3), nrow = 2, ncol = 2, byrow = FALSE))
boxplot(I(width * convert) ~ unit, data = roomwidth, ylab = "Estimated width (feet)", varwidth = TRUE,
names = c("Estimates in feet", "Estimates in metres (converted to feet)"))
feet <- roomwidth$unit == "feet"
metres <- roomwidth$unit == "metres"
qqnorm(roomwidth$width[feet], ylab = "Estimated width (feet)")
qqline(roomwidth$width[feet])
qqnorm(roomwidth$width[metres], ylab = "Estimated width (metres)")
qqline(roomwidth$width[metres])
```

The boxplots indicate that both sets of estimates contain a number of outliers and also that the estimates made
in metres are skewed and more variable than those made in feet, a point underlined by the numerical summary statistics above. Both normal probability plots depart from linearity, suggesting that the distributions of both sets of estimates are not normal. The presence of outliers, the apparently different variances and the evidence of non-normality all suggest caution in applying the t-test, but for the moment we shall apply the usual version of the test using the t.test function in 

Independent samples t-test
```{r rw_tt1}
t.test(I(width * convert) ~ unit, data = roomwidth, var.equal = TRUE)
```

Independent samples Welch test (notice var.equal parameter set to FALSE)
```{r rw_tt2}
t.test(I(width * convert) ~ unit, data = roomwidth, var.equal = FALSE)
```

Wilcoxon rank sum test (non-parametric) for shaky assumptions case

```{r rw_np_wilcox}
wilcox.test(I(width * convert) ~ unit, data = roomwidth, conf.int = TRUE)
```


## wave Energy Device Mooring
```{r we_getready, echo=FALSE}
data("waves", package = "HSAUR")
```
```{r we_head}
waves[1:4,]
```
This data is Measurements of root mean square bending moment by two different mooring methods and requires matched pairs t-test to answer the question of interest. This test assumes that the differences between the matched observations
have a normal distribution so we can begin by checking this assumption by
constructing a boxplot and a normal probability plot

```{r we_plots}
mooringdiff <- waves$method1 - waves$method2
layout(matrix(1:2, ncol = 2))
boxplot(mooringdiff, ylab = "Differences (Newton metres)", main = "Boxplot")
abline(h = 0, lty = 2)
qqnorm(mooringdiff, ylab = "Differences (Newton metres)")
qqline(mooringdiff)
```

paired t-test for the waves data
```{r we_tt1}
t.test(mooringdiff)
```

Wilcoxon signed rank test
```{r we wt1}
wilcox.test(mooringdiff)
```



## Mortality and Water Hardness
```{r water, echo=FALSE}
data("water", package = "HSAUR")
```
```{r water_show}
head(water)
```

This data is the mortality and drinking water hardness for 61 cities in England and Wales.

```{r water_plot}
nf <- layout(matrix(c(2, 0, 1, 3), 2, 2, byrow = TRUE), c(2, 1), c(1, 2), TRUE)
psymb <- as.numeric(water$location)
plot(mortality ~ hardness, data = water, pch = psymb)
abline(lm(mortality ~ hardness, data = water))
legend("topright", legend = levels(water$location), pch = c(1, 2), bty = "n")
hist(water$hardness)
boxplot(water$mortality)
```

Hardness shows it has a rather skewed distribution.
Enhanced scatterplot of water hardness and mortality, showing both the joint and the marginal distributions and, in addition, the location of the city by different plotting symbols.

The scatterplot shows that as hardness increases mortality decreases, and the histogram for the water hardness shows it has a rather skewed distribution.

```{r water_corr}
cor.test(~mortality + hardness, data = water)
```




## Piston-ring Failures
```{r pr, echo=FALSE}
data("pistonrings", package = "HSAUR")
```
```{r pr_show}
head(pistonrings)
```
Data describes number of failures of piston rings in three legs of four steam-driven compressors

Rather than looking at the simple differences of observed and expected values for each cell which would be unsatisfactory since a difference of fixed size is clearly more important for smaller samples, it is preferable to consider a standardised residual given by dividing the observed minus expected difference by the square root of the appropriate expected value. The X^2 statistic for
assessing independence is simply the sum, over all the cells in the table, of the squares of these terms. We can find these values extracting the residuals element of the object returned by the chisq.test function.

```{r pr_xsr}
chisq.test(pistonrings)$residuals
```

Association plot of the residuals
```{r pr_xsr_plot}
library("vcd")
assoc(pistonrings)
```

Standard Chi-squared test
```{r pr_xs}
chisq.test(pistonrings)
```



## Rearrests of Juveniles
```{r jv, echo=FALSE}
data("rearrests", package = "HSAUR")
```
```{r jv_show}
head(rearrests)
```

The data shows rearrests of juventile felons by type of court in which they were tried.

```{r jv_mt}
mcnemar.test(rearrests, correct = FALSE)
```

Here we need to use McNemar’s test to assess whether rearrest is associated with type of court where the juvenile was tried. We can use the R function mcnemar.test. The test statistic is 62.888 with a single degree of freedom – the associated p-value is extremely small and there is strong evidence that type of court and the probability of rearrest are related. It appears that trial at a juvenile court is less likely to result in rearrest

```{r jv_mc}
mcnemar.test(rearrests, correct = FALSE)
```
```{r jv_bn}
binom.test(rearrests[2], n = sum(rearrests[c(2,3)]))
```

*Exact version of McNemar’s test for the rearrests data computed via a binomal test.*


